---
title: "compare_nldas"
output: 
  html_document:
   keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(janitor)

```

Compare `nldas_8` temperatures (data rod) with prior dataset

```{r}
nldas_8 <-
  read_rds("output/nldas_8.Rmd")

daily_indices_old <-
  read_rds("C:/Users/slewa/OneDrive - cumc.columbia.edu/daily_hsi/data/daily_indices.rds") %>% 
  dplyr::select(c(installation, date, tmp_f_mean, tmp_f_max))

```

Create daily indices (mean, max daily temp) from `nldas_8`
```{r}

nldas8_daily <-
  nldas_8 %>% 
    unnest() %>% 
    dplyr::select(site_name, datetime, TMP2m) %>% 
    group_by(site_name, lubridate::date(datetime)) %>% 
    summarise(tmp_mean = mean(TMP2m),
              tmp_max = max(TMP2m)) %>% 
    mutate(tmp_mean = (tmp_mean -273.15) * 9/5 + 32,
           tmp_max = (tmp_max -273.15) * 9/5 + 32) %>% 
  rename(date = `lubridate::date(datetime)`)

```

Prepare old df to join
```{r}


daily_indices_old <-
  daily_indices_old %>% 
    mutate(installation = recode(installation,
                                 eglin_afb = "Eglin AFB",
                                  fort_benning_ga = "Fort Benning",
                                  fort_bliss = "Fort Bliss",
                                  fort_bragg = "Fort Bragg",
                                  fort_campbell = "Fort Campbell",
                                  fort_gordon = "Fort Gordon",
                                  fort_hood = "Fort Hood",
                                  fort_jackson = "Fort Jackson",
                                  fort_knox = "Fort Knox",
                                  fort_leonard_wood = "Fort Leonard Wood",
                                  fort_polk = "Fort Polk",
                                  fort_riley = "Fort Riley",
                                  fort_sam_houston = "Joint Base San Antonio",
                                  fort_sill = "Fort Sill",
                                  fort_stewart = "Fort Stewart",
                                  mcb_camp_lejeune = "MCB Camp Lejeune",
                                  mcb_camp_pendleton = "MCB Camp Pendleton",
                                  mcb_quantico = "MCB Quantico",
                                  mcrd_beaufort_parris_island = "MCRD Parris Island",
                                  mcrd_san_diego = "MCRD San Diego",
                                  ntc_and_fort_irwin = "Fort Irwin",
                                  pensacola = "NAS Pensacola",
                                  portsmouth = "NMC Portsmouth",
                                  twentynine_palms_main_base = "Twentynine Palms")) %>% 
    rename(tmp_mean_old = tmp_f_mean,
           tmp_max_old = tmp_f_max)


```

# Join "old" to "new" (data rods)

```{r}
joined_daily_temp <-
  nldas8_daily %>% 
    left_join(daily_indices_old, by = c("site_name" = "installation", "date" = "date"))


```

# Inspect differences
Note: times averaged in new set from UTC, not local


Plots
```{r}
# 1996 - 2019
joined_daily_temp %>% 
  filter(lubridate::year(date) > 1995 & lubridate::year(date) < 2020) %>% 
  ggplot(aes(x = date)) +
    geom_line(aes(y = tmp_mean), color = "blue") +
    geom_line(aes(y = tmp_mean_old), color = "red") +
    facet_wrap(~ site_name)

# July 2000

joined_daily_temp %>% 
  filter(lubridate::year(date) == 2000,
         lubridate::month(date) == 7) %>% 
  ggplot(aes(x = date)) +
    geom_line(aes(y = tmp_mean), color = "blue") +
    geom_line(aes(y = tmp_mean_old), color = "red") +
    facet_wrap(~ site_name)

#ggsave("output/compare_daily_mean_tmp_jul2000.png", width = 10, height = 10)  


joined_daily_temp %>% 
  filter(lubridate::year(date) == 2000,
         lubridate::month(date) == 7) %>% 
  ggplot(aes(x = date)) +
    geom_line(aes(y = tmp_max), color = "blue") +
    geom_line(aes(y = tmp_max_old), color = "red") +
    facet_wrap(~ site_name)
#ggsave("output/compare_daily_max_tmp_jul2000.png", width = 10, height = 10)  
```

Good: eglin, benning, knox, jbsa, lejeune, pendleton, quantico, parris island, san diego, pensacola, portsmouth, 29 palms (12)
Off: bliss, bragg, campbell, gordon, hood, irwin, jackson, leonard wood, polk, riley, sill, stewart



# Paired t-test
```{r}
joined_daily_temp %>%
  dplyr::select(-tmp_max, -tmp_max_old) %>% 
  gather(key = variable, value = value, -c(site_name, date)) %>% 
  group_by(site_name, variable) %>%
  summarise(value = list(value)) %>%
  spread(variable, value) %>%
  group_by(site_name) %>% 
      mutate(p_value = t.test(unlist(tmp_mean), unlist(tmp_mean_old), paired = TRUE)$p.value,
         t_value = t.test(unlist(tmp_mean), unlist(tmp_mean_old), paired = TRUE)$statistic) 

```


