---
title: "gldasR_query_pressure"
output: 
  html_document:
   keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(janitor)
library(sf)
library(remotes)
library(nldastools) #remotes::install_github("lawinslow/nldasR", force = TRUE)
library(httr)

```

## GLDAS 3 hour Data rod query

Global Land Data Assimilation System
GLDAS Noah Land Surface Model L4 3 hourly 0.25 x 0.25 degree V2.1 (GLDAS_NOAH025_3H)
https://disc.gsfc.nasa.gov/datasets/GLDAS_NOAH025_3H_2.1/summary  



https://rdrr.io/github/lawinslow/nldasR/man/nldas2_primary_forcing_rods.html 
Interfaces with NASA Data Rods (Time series) services managed by Goddard Earth Sciences Data and Information Services Center (GES DISC). More info about the interface available here.


```{r}

centroids <-
  read_rds("output/centroids.Rmd")

centroids %>% as.data.frame()

```

  https://hydro1.gesdisc.eosdis.nasa.gov/daac-bin/access/timeseries.cgi?variable=GLDAS2:GLDAS_NOAH025_3H_v2.1:Psurf_f_inst&startDate=2002-07-01T00&endDate=2002-07-03T00&location=GEOM:POINT(-84.8012, 32.3996)&type=asc2

## Get GLDAS function:
Adapted from: https://rdrr.io/github/lawinslow/nldasR/src/R/nldas2_primary_forcing_rods.R

```{r}

#' @title Get GLDAS primary forcing data
#'
#' @param lon Longitude of data location request
#' @param lat Latitude of data location request
#' @param start Start date as POSIXct
#' @param stop End date as POSIXct
#'
#' @description
#' Interfaces with NASA Data Rods (Time series) services managed by Goddard Earth Sciences Data and
#' Information Services Center (GES DISC). More info about the interface available
#' \href{https://disc.sci.gsfc.nasa.gov/hydrology/data-rods-time-series-data}{here}.
#'
#' @import httr
#' @export
gldas2_1_rods <- function(lon, lat, start=as.POSIXct('2000-01-01', tz='UTC'), stop=as.POSIXct('2017-01-01', tz='UTC')){

  if(length(lon) != 1 || length(lat) != 1){
    stop('Lat and Lon must both be length of 1')
  }

  if(start < as.POSIXct('1999-12-31', tz='UTC') || start > as.POSIXct(Sys.Date())){
    stop('Start date must be between 1999-12-31 and now')
  }

  if(stop < as.POSIXct('1999-12-31', tz='UTC') || stop > as.POSIXct(Sys.Date())){
    stop('Stop date must be between 1999-12-31 and now')
  }

  pf_vars = "Psurf_f_inst"


  url_pattern = paste0('https://hydro1.gesdisc.eosdis.nasa.gov/daac-bin/access/timeseries.cgi?',
                       'variable=GLDAS2:GLDAS_NOAH025_3H_v2.1:%s&location=GEOM:POINT(%g,%%20%g)&startDate=%s&endDate=%s&type=asc2')


  dl_parse = function(var){
    durl = sprintf(url_pattern, var, lon, lat, format(start, '%Y-%m-%dT%H'), format(stop, '%Y-%m-%dT%H'))
    r = RETRY('GET', durl)
    d = read.table(text=content(r, 'text'), skip=12, header=TRUE, fill=TRUE)
    names(d) = c('datetime', var)
    d$datetime = as.POSIXct(strptime(paste0(d$datetime, 'T', d$hour), '%Y-%m-%dT%H', tz='UTC'))
    d$hour = NULL
    return(d)
  }

  d_list = lapply(pf_vars, dl_parse)

  pf_data = Reduce(function(...){merge(..., by='datetime')}, d_list)
  return(pf_data)
}
```




## Test interface
Source code: https://rdrr.io/github/lawinslow/nldasR/src/R/nldas2_primary_forcing_rods.R



https://hydro1.gesdisc.eosdis.nasa.gov/daac-bin/access/timeseries.cgi?variable=GLDAS2:GLDAS_NOAH025_3H_v2.1:Psurf_f_inst&startDate=2002-07-01T00&endDate=2002-07-03T00&location=GEOM:POINT(-84.8012, 32.3996)&type=asc2



```{r}
#short
gldas2_1_rods(-84.80115, 32.39959, start = as.POSIXct("2002-07-01", tz =
  "UTC"), stop = as.POSIXct("2002-07-03", tz = "UTC"))

# start date test

gldas2_1_rods(-84.80115, 32.39959, start = as.POSIXct("2000-01-01", tz =
  "UTC"), stop = as.POSIXct("2000-01-03", tz = "UTC"))

# end date test
gldas2_1_rods(-84.80115, 32.39959, start = as.POSIXct("2021-12-28", tz =
  "UTC"), stop = as.POSIXct("2021-12-31", tz = "UTC"))


centroids %>% as.data.frame()

```


## Get GLDAS 2.1 surface pressure rods (2000-2022)

```{r}

start_date <- as.POSIXct("2000-01-01", tz =  "UTC")
stop_date <-  as.POSIXct("2022-01-01", tz = "UTC")



gldas_2_1_press_00_22 <-
  centroids %>%
      mutate(gldas = map2(.x = lon, .y = lat, 
                          ~gldas2_1_rods(.x, .y, start = start_date, stop = stop_date)))
#write_rds(gldas_2_1_press_00_22, "output/gldas_2_1_press_00_22")

gldas_2_1_press_00_22

```


## Function to get GLDAS2.0 (1948-2014)

https://hydro1.gesdisc.eosdis.nasa.gov/daac-bin/access/timeseries.cgi?variable=GLDAS2:GLDAS_NOAH025_3H_v2.0:Psurf_f_inst&startDate=2002-07-01T00&endDate=2002-07-03T00&location=GEOM:POINT(-84.8012, 32.3996)&type=asc2


```{r}

#' @title Get GLDAS primary forcing data
#'
#' @param lon Longitude of data location request
#' @param lat Latitude of data location request
#' @param start Start date as POSIXct
#' @param stop End date as POSIXct
#'
#' @description
#' Interfaces with NASA Data Rods (Time series) services managed by Goddard Earth Sciences Data and
#' Information Services Center (GES DISC). More info about the interface available
#' \href{https://disc.sci.gsfc.nasa.gov/hydrology/data-rods-time-series-data}{here}.
#'
#' @import httr
#' @export
gldas2_0_rods <- function(lon, lat, start=as.POSIXct('2000-01-01', tz='UTC'), stop=as.POSIXct('2017-01-01', tz='UTC')){

  if(length(lon) != 1 || length(lat) != 1){
    stop('Lat and Lon must both be length of 1')
  }

  if(start < as.POSIXct('1948-12-31', tz='UTC') || start > as.POSIXct(Sys.Date())){
    stop('Start date must be between 1948-12-31 and now')
  }

  if(stop < as.POSIXct('1948-12-31', tz='UTC') || stop > as.POSIXct(Sys.Date())){
    stop('Stop date must be between 1948-12-31 and now')
  }

  pf_vars = "Psurf_f_inst"


  url_pattern = paste0('https://hydro1.gesdisc.eosdis.nasa.gov/daac-bin/access/timeseries.cgi?',
                       'variable=GLDAS2:GLDAS_NOAH025_3H_v2.0:%s&location=GEOM:POINT(%g,%%20%g)&startDate=%s&endDate=%s&type=asc2')


  dl_parse = function(var){
    durl = sprintf(url_pattern, var, lon, lat, format(start, '%Y-%m-%dT%H'), format(stop, '%Y-%m-%dT%H'))
    r = RETRY('GET', durl)
    d = read.table(text=content(r, 'text'), skip=12, header=TRUE, fill=TRUE)
    names(d) = c('datetime', var)
    d$datetime = as.POSIXct(strptime(paste0(d$datetime, 'T', d$hour), '%Y-%m-%dT%H', tz='UTC'))
    d$hour = NULL
    return(d)
  }

  d_list = lapply(pf_vars, dl_parse)

  pf_data = Reduce(function(...){merge(..., by='datetime')}, d_list)
  return(pf_data)
}
```



## Get GLDAS 2.0 surface pressure rods (1990-2000)

```{r}

start_date <- as.POSIXct("1989-12-31", tz =  "UTC")
stop_date <-  as.POSIXct("2000-01-01", tz = "UTC")



gldas_2_0_press_00_22 <-
  centroids %>%
      mutate(gldas = map2(.x = lon, .y = lat, 
                          ~gldas2_0_rods(.x, .y, start = start_date, stop = stop_date)))
#write_rds(gldas_2_0_press_00_22, "output/gldas_2_0_press_00_22")

gldas_2_0_press_00_22  #1990-2000

```


## Join GLDAS 1990-2000 (2.0) with 2000-2022 (2.1)
```{r}

gldas_pressure <-
gldas_2_0_press_00_22 %>% 
  bind_rows(gldas_2_1_press_00_22) %>% 
  unnest(nldas) %>% 
  group_by(site_name) %>% 
  arrange(site_name, datetime) %>% 
  rename(pressure = Psurf_f_inst)


gldas_pressure %>% 
  summary()
```


