---
title: "nldasR_query"
output: 
  html_document:
   keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(janitor)
library(sf)
library(remotes)
library(nldastools)
#remotes::install_github("lawinslow/nldasR", force = TRUE)

```

## NLDAS 2 Data rod query

https://rdrr.io/github/lawinslow/nldasR/man/nldas2_primary_forcing_rods.html 
Interfaces with NASA Data Rods (Time series) services managed by Goddard Earth Sciences Data and Information Services Center (GES DISC). More info about the interface available here.

Provides: datetime APCPsfc DLWRFsfc DSWRFsfc PEVAPsfc   SPFH2m  TMP2m UGRD10m VGRD10m
Surface Pressure band not available through this data server :(


```{r}

centroids <-
  read_rds("output/centroids.Rmd")

centroids

```


## Test interface
Source code: https://rdrr.io/github/lawinslow/nldasR/src/R/nldas2_primary_forcing_rods.R

```{r}
#short
nldas2_primary_forcing_rods(-84.80115, 32.39959, start = as.POSIXct("1990-07-14", tz =
  "UTC"), stop = as.POSIXct("1990-07-15", tz = "UTC"))

#long
nldas2_primary_forcing_rods(-84.80115, 32.39959, start = start_date, stop = stop_date)

# getting error for Pensacola: 
nldas2_primary_forcing_rods(-87.30628, 30.35261, start = start_date, stop = stop_date)

centroids %>% as.data.frame()

nldas2_primary_forcing_rods(-117.19769, 32.74017, start = as.POSIXct("1990-07-14", tz =
  "UTC"), stop = as.POSIXct("1990-07-15", tz = "UTC"))
```


## Get NLDAS 2 primary forcing rods

```{r}

start_date <- as.POSIXct("1989-12-31", tz =  "UTC")
stop_date <-  as.POSIXct("2022-01-01", tz = "UTC")


# Times out with 24 rows "Request failed [404]"; break into smaller slices
# 3 coordinate sets return error: Pensacola, San Diego, Parris Island -- add 0.05 degree to lat and lon


nldas_8band_a <-
  centroids %>%
      slice(1:2) %>% 
      mutate(nldas = map2(.x = lon, .y = lat, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))
#write_rds(nldas_8band_a, "output/nldas_8band_a")

nldas_8band_b1 <-
  centroids %>%
      slice(3) %>%  # Pensacola
      mutate(nldas = map2(.x = lon + 0.05, .y = lat + 0.05, # adjustment needed to return data
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))

#write_rds(nldas_8band_b1, "output/nldas_8band_b1")

nldas_8band_b2 <-
  centroids %>%
      slice(4) %>% 
      mutate(nldas = map2(.x = lon, .y = lat, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))
#write_rds(nldas_8band_b2, "output/nldas_8band_b2")

nldas_8band_b3 <-
  centroids %>%
      slice(5) %>% 
      mutate(nldas = map2(.x = lon, .y = lat, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))
#write_rds(nldas_8band_b3, "output/nldas_8band_b3")

nldas_8band_b4 <-
  centroids %>%
      slice(6) %>% # San Diego
      mutate(nldas = map2(.x = lon + 0.05, .y = lat + 0.05, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))
#write_rds(nldas_8band_b4, "output/nldas_8band_b4")

nldas_8band_c <-
  centroids %>%
      slice(7:9) %>% 
      mutate(nldas = map2(.x = lon, .y = lat, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))
# write_rds(nldas_8band_c, "output/nldas_8band_c")

nldas_8band_d <-
  centroids %>%
      slice(10:11) %>% 
      mutate(nldas = map2(.x = lon, .y = lat, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))
# write_rds(nldas_8band_d, "output/nldas_8band_d")

nldas_8band_12 <-
  centroids %>%
      slice(12) %>% 
      mutate(nldas = map2(.x = lon, .y = lat, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))

# write_rds(nldas_8band_12, "output/nldas_8band_12")

nldas_8band_13 <-
  centroids %>%
      slice(13) %>% # Parris Island
      mutate(nldas = map2(.x = lon + 0.05, .y = lat + 0.05, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))

# write_rds(nldas_8band_13, "output/nldas_8band_13")

nldas_8band_14_15 <-
  centroids %>%
      slice(14:15) %>% 
      mutate(nldas = map2(.x = lon, .y = lat, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))

# write_rds(nldas_8band_14_15, "output/nldas_8band_14_15")

nldas_8band_16_17 <-
  centroids %>%
      slice(16:17) %>% 
      mutate(nldas = map2(.x = lon, .y = lat, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))

# write_rds(nldas_8band_16_17, "output/nldas_8band_16_17")

nldas_8band_18_19 <-
  centroids %>%
      slice(18:19) %>% 
      mutate(nldas = map2(.x = lon, .y = lat, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))

# write_rds(nldas_8band_18_19, "output/nldas_8band_18_19")

nldas_8band_20_22 <-
  centroids %>%
      slice(20:22) %>% 
      mutate(nldas = map2(.x = lon, .y = lat, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))
# write_rds(nldas_8band_20_22, "output/nldas_8band_20_22") 

nldas_8band_23_24 <-
  centroids %>%
      slice(23:24) %>% 
      mutate(nldas = map2(.x = lon, .y = lat, 
                          ~nldas2_primary_forcing_rods(.x, .y, start = start_date, stop = stop_date)))

# write_rds(nldas_8band_23_24, "output/nldas_8band_23_24") 

```

## Compile into single dataframe

```{r}


```























get_primary_forcing_local_multitool(lons, lats, ids)

















