---
title: "nldas_df"
output: 
  html_document:
   keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(janitor)
library(lutz)
library(sf)
library(weathermetrics)
library(humidity)
library(inspectdf)
#devtools::install_github("mdljts/wbgt")
```

Input: "output/nldas8.rds", "output/nldas_pressure.rds", "data/centroid_coords.rds" (for time zones)
Output: "data/nldas_hourly_rods.rds", "index_hourly_rod.rds", "output/time_zones.rds"


Prepare hourly NLDAS2 df and hourly index dataframes

```{r}
nldas_8 <-
  read_rds("output/nldas_8")

nldas_pressure <-
  read_rds("output/pressure_8.rds")

```


## Join surface pressure to NLDAS 2 bands
```{r}

nldas_hourly_rods <- #6,732,888 rows x 11 columns
  nldas_8 %>% 
    dplyr::select(-c(lon, lat)) %>% 
    unnest(nldas) %>% 
    left_join(nldas_pressure, by = c("site_name" = "name", "datetime" = "timestamp"))
  

```

## Local time from UTC

Timezones for local time conversion
```{r}
centroids <- read_rds(file = "data/centroid_coords.rds")

time_zones <-
  centroids %>% 
    dplyr::filter(!site_name %in% c("Fort Huachuca","Fort Carson","Fort Lewis", "Lackland AFB",
                                    "West Point Mil Reservation","Fort Drum")) %>%
    mutate(site_name = recode(site_name, `Fort Benning GA` = "Fort Benning",
                              `Naval Medical Center Portsmouth` = "NMC Portsmouth",
                              `NTC and Fort Irwin` = "Fort Irwin",
                              `Twentynine Palms Main Base` = "Twentynine Palms",
                              `MCRD Beaufort Parris Island` = "MCRD Parris Island",
                              `Fort Sam Houston` = "Joint Base San Antonio")) %>% 
  dplyr::select(site_name, centroid) %>% 
  mutate(time_zone = lutz::tz_lookup(centroid, crs = 4326, method = "accurate")) %>% 
  as_tibble() 

# write_rds(time_zones, "output/time_zones.rds")
time_zones <-
  read_rds("output/time_zones.rds")


time_zones %>% as.data.frame() 

# add local dttm

nldas_hourly_rods <-
  nldas_hourly_rods %>% 
    rename(utc_dttm = datetime) %>% 
    left_join(time_zones %>% dplyr::select(-centroid), by = "site_name") %>%
    dplyr::mutate(local_dttm = dplyr::case_when(
        .$time_zone == "America/New_York" ~ as.character(with_tz(.$utc_dttm, tz = "America/New_York")),
        .$time_zone == "America/Chicago" ~ as.character(with_tz(.$utc_dttm, tz = "America/Chicago")), 
        .$time_zone == "America/Denver" ~ as.character(with_tz(.$utc_dttm, tz = "America/Denver")), 
        .$time_zone == "America/Los_Angeles" ~ as.character(with_tz(.$utc_dttm, tz = "America/Los_Angeles")))) 


nldas_hourly_rods <-
  nldas_hourly_rods %>% 
    dplyr::select(-time_zone) %>% 
    dplyr::select(site_name, utc_dttm, local_dttm, everything())

# write_rds(nldas_hourly_rods, "data/nldas_hourly_rods.rds")   # 6,732,888 rows x 12 columns

nldas_hourly_rods <-
  read_rds("data/nldas_hourly_rods.rds")
  

```

## Check mean temperatures

```{r}

nldas_hourly_rods %>% 
  group_by(site_name) %>% 
  summarise(mat = 1.8 * (mean(TMP2m, na.rm = TRUE)-273.15) + 32) %>% 
  as.data.frame()

```

## Meteorology conversions and heat index
From original NLDAS variables, add windspeed, temp in F and C, relative humidity, dew point (C), heat index (F)

```{r}

nldas_met_conversions <-
  nldas_hourly_rods %>%
    dplyr::filter(lubridate::date(utc_dttm) > "1989-12-31" & lubridate::date(utc_dttm) < "2020-01-01") %>% 
    dplyr::select(-c(APCPsfc, PEVAPsfc)) %>% 
    mutate(windspeed = sqrt(UGRD10m^2 + VGRD10m^2),
             tmp_f = weathermetrics::convert_temperature(TMP2m, old_metric = "kelvin", new_metric = "fahrenheit"),
             tmp_c = weathermetrics::convert_temperature(TMP2m, old_metric = "kelvin", new_metric = "celsius"),
             rh = humidity::SH2RH(SPFH2m, TMP2m, p = pressure, isK = TRUE),
             rh = case_when(
               rh > 100 ~ 100,
               rh <= 100 ~ rh
             ), # Restrict RH to 100% maximum
             dp_c = humidity.to.dewpoint(rh = rh, t = tmp_c, temperature.metric = "celsius"),
             heat_index = weathermetrics::heat.index(t = tmp_f,
                             rh = rh,
                             temperature.metric = "fahrenheit",
                             output.metric = "fahrenheit",
                             round = 2)) %>%
    dplyr::select(-c(UGRD10m, VGRD10m))
  
nldas_met_conversions 

nldas_met_conversions %>% summary()

nldas_met_conversions %>% 
  inspectdf::inspect_na() # no missing values


```

## Wet Bulb Globe Temperature (WBGT) Estimate

- apply a wrapper of the C code developed by Liljegren et al to estimate WBGT from datasets of standard meterological measurements; product  produced by UChicago Argonne, LLC under contract with the Department of Energy

- need lat and lon columns for the WBGT function


### Apply lon/lat, calculate local - utc time difference, create dataframe with columns needed for wbgt::wbgt_df (mdljts wrapper) 
```{r}

lon_lat <-
  time_zones %>%
    mutate(longitude = unlist(map(centroid, 1)),
           latitude = unlist(map(centroid, 2))) %>% 
    dplyr::select(-centroid)
    

# Create datframe for `wbgt_df` wrapper function

  # LST - GMT difference (utc_offset)
to_wbgt_prep <-  
  nldas_met_conversions %>% 
      left_join(lon_lat, by = "site_name") %>% 
      mutate(local_dttm = lubridate::as_datetime(local_dttm),
             utc_offset = lubridate::as.period(local_dttm - utc_dttm, unit = hours)) %>% # hour difference in period format
      mutate(utc_offset = lubridate::time_length(utc_offset, unit = "hours")) # hour difference in numeric format

  # function input dataframe           
to_wbgt <-  
  to_wbgt_prep %>%  
     mutate(year = lubridate::year(local_dttm), # 4-digit, e.g. 2007	
            month = lubridate::month(local_dttm), # month (1-12)
            day = lubridate::day(local_dttm), # day of month
            hour = lubridate::hour(local_dttm), # hour in local standard time (LST)
            minute = lubridate::minute(local_dttm), # minutes past the hour	
            gmt = utc_offset, # LST-GMT difference, hours (negative in USA)	
            avg = 60, #  averaging time of meteorological inputs, minutes
            lat = latitude, # north latitude, decimal		
            lon = longitude, # east longitude, decimal (negative in USA)
            solar = DSWRFsfc,  # solar irradiance, W/m2
            pres = pressure / 100, # barometric pressure, mb		
            Tair = tmp_c, # air (dry bulb) temperature, degC	
            relhum = rh, # relative humidity, %	
            speed = windspeed, # wind speed, m/s
            zspeed = 10, # height of wind speed measurement, m	
            dT = 0, #  vertical temperature difference (upper minus lower), degC
            urban = 1) %>% # urban best suited for semi-woodland (Liljegren)
    dplyr::select(year:urban)



# Execute WBGT wrapper on entire dataframe 
wbgt_mdljts <-     
 wbgt::wbgt_df(to_wbgt) 

# write_rds(wbgt_mdljts, "data/wbgt_mdljts_output.rds")

wbgt_mdljts <- 
  read_rds("data/wbgt_mdljts_output.rds")

# Bind WBGT column to NLDAS dataframe and add Fahrenheit conversion

  # select WBGT column (deg C)
wbgt_mdljts <- 
  wbgt_mdljts %>%
    dplyr::select(wbgt)
  
  # bind wbgt, corresponding 6,111,232 row dataframes

nldas_hourly_index_full <-
  nldas_met_conversions %>% 
  bind_cols(wbgt_mdljts) %>% 
  mutate(wbgt_f = weathermetrics::convert_temperature(wbgt, old_metric = "celsius", new_metric = "fahrenheit")) %>% 
  rename(wbgt_c = wbgt,
         hi_f = heat_index)

# write_rds(nldas_hourly_index_full, "data/nldas_hourly_index_full.rds")
    

nldas_hourly_index_rod <-
  nldas_hourly_index_full %>% 
    dplyr::select(-c(DLWRFsfc, DSWRFsfc, SPFH2m, TMP2m, pressure, windspeed, rh, dp_c)) %>%
    mutate(hi_c = weathermetrics::convert_temperature(hi_f, old_metric = "fahrenheit", new_metric = "celsius")) 

# write_rds(nldas_hourly_index_rod, "data/nldas_hourly_index_rod.rds")  
```



