---
title: "surf_press_from_gee"
output: 
  html_document:
   keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(janitor)
library(googledrive)
library(httpuv)
library(Rcpp)
library(stringr)

```

Import NLDAS 2 surface pressure CSV files from Google Drive, reshape, compile batches, and join with other NLDAS bands.
Values extracted through GEE script in `get_surface_pressure_earth_engine`.
Reshape from wide (timestamp as columns) to long (timestamp as variable).

Files are hourly by decade (task runtime from 3 - 10 hours each). A 32-year task timed out after 12 hours.


## Download files from Google Drive
```{r}
# Access via Tidyverse `googledrive` package

drive_auth() # complete authorization steps/permissions on browser

drive_find(n_max = 30)

drive_download("~/pressure_decade/pressure1990.csv", path = "data/pressure1990.csv")
drive_download("~/pressure_decade/pressure2000.csv", path = "data/pressure2000.csv")
drive_download("~/pressure_decade/pressure2010.csv", path = "data/pressure2010.csv")

```


## Read and clean NLDAS 2 surface pressure files
1st list (19 sites)
```{r}

pressure1990 <-
  read_csv(file = "data/pressure1990.csv")

pressure1990 # 19 rows, 87,651 columns
pressure1990$name  # missing: Hood, Bliss, Leonard Wood, Knox, Stewart; missing from geometry point features list

pressure1990_l <- #1,665,312 rows x 3 columns
  pressure1990 %>% 
    dplyr::select(name, `.geo`, `system:index`, everything()) %>% 
    dplyr::select(-c(`.geo`, `system:index`)) %>% 
    pivot_longer(-name, names_to = "timestamp", values_to = "pressure")


pressure1990_l %>% 
  tail()  # through A19991231_2300

# 2000 - 2010
pressure2000 <-
  read_csv(file = "data/pressure2000.csv")

pressure2000 # 19 rows, 87,651 columns
pressure2000$name  # missing: Hood, Bliss, Leonard Wood, Knox, Stewart; missing from geometry point features list

pressure2000_l <- #1,665,312 rows x 3 columns
  pressure2000 %>% 
    dplyr::select(name, `.geo`, `system:index`, everything()) %>% 
    dplyr::select(-c(`.geo`, `system:index`)) %>% 
    pivot_longer(-name, names_to = "timestamp", values_to = "pressure")

# start: A20000101_0000
# end:
pressure2000_l %>% 
  tail()  # through A20091231_2300

#2010 - 2020


pressure2010 <-
  read_csv(file = "data/pressure2010.csv")

pressure2010 # 19 rows, 87,651 columns
pressure2010$name  # missing: Hood, Bliss, Leonard Wood, Knox, Stewart; missing from geometry point features list

pressure2010_l <- #1,665,312 rows x 3 columns
  pressure2010 %>% 
    dplyr::select(name, `.geo`, `system:index`, everything()) %>% 
    dplyr::select(-c(`.geo`, `system:index`)) %>% 
    pivot_longer(-name, names_to = "timestamp", values_to = "pressure")

# start: A20100101_0000
# end:
pressure2010_l %>% 
  tail()  # through A20191231_2300


pressure_long_a <-
  pressure1990_l %>% 
    bind_rows(pressure2000_l) %>% 
    bind_rows(pressure2010_l)


pressure_long_a
# write_rds(pressure_long_a, "data/pressure_long_a.rds")  
# 4,996,392 x 3, NLDAS2 surface pressure for 19 sites

```


## Read and clean NLDAS 2 surface pressure files
2nd list (5 sites)
Executed in 1 run, 1990 - 2020 in 4 hours with bounding box applied to NLDAS image collection stack

```{r}

drive_download("~/pressure_5sites/pressureb_all.csv", path = "data/pressureb_all.csv")


pressureb <-
  read_csv(file = "data/pressureb_all.csv")

pressureb # 19 rows, 262,995 columns
pressureb$name   # Hood, Bliss, Leonard Wood, Knox, Stewart

# Wide to Long
pressureb_l <- #1,314,960 rows x 3 columns
  pressureb %>% 
    dplyr::select(name, `.geo`, `system:index`, everything()) %>% 
    dplyr::select(-c(`.geo`, `system:index`)) %>% 
    pivot_longer(-name, names_to = "timestamp", values_to = "pressure")


# write_rds(pressureb_l, "data/pressure_long_b.rds") 

```


## Bind long dataframes (19 sites and 5 sites) of surface pressure data

```{r}
pressurea_l <-
  read_rds("data/pressure_long_a.rds")


pressurea_l %>% 
  slice(1:3) %>% 
    mutate(timestamp = str_sub(timestamp, 2, -1),
           timestamp = lubridate::as_datetime(timestamp,
                                              format = "%Y%m%d_%H%M", 
                                              tz = "UTC")) 
    


nldas_pressure <-
  pressurea_l %>% 
  bind_rows(pressureb_l) %>% 
  mutate(timestamp = str_sub(timestamp, 2, -1),
           timestamp = lubridate::as_datetime(timestamp,
                                              format = "%Y%m%d_%H%M", 
                                              tz = "UTC"))

# write_rds(nldas_pressure, "output/nldas_pressure.rds")  # 3 columns x 6,311,352 rows
```

